Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory

### 2016-10-06 KKosmology.tex
After the comments by R. Olea

The simplest possible extra term in the Lanczos--Lovelock action  is a quadratic construction of curvatures, called Gauss--Bonnet term, which reads
\begin{equation}\label{GB}
  \mathcal{L}_{\mathrm{GB}} = \dn{N}{x}\sqrt{-g}\left(\tilde{R}^2-4\tilde{R}_{\mu\nu}\tilde{R}^{\mu\nu}+\tilde{R}_{\alpha\beta\mu\nu}
  \tilde{R}^{\alpha\beta\mu\nu}\right),
\end{equation}
where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant. $\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively. In four dimensions, the Gauss--Bonnet term adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term. Nevertheless, in dimensions higher than four it contributes to the field equations. This Lagrangian was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.

@@

In this work, we will focus on the region in the parameter space where
\begin{equation}
  \label{delta}
  \Delta\equiv\alpha_1^2-4\alpha_0\alpha_2\neq 0.
\end{equation}
This condition place us outside the Chern--Simons point. Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2016cs,*Troncoso:1999pk}. Exact solutions with torsion in Chern--Simons gravity are given in~\cite{Aros:2007nn,Banados:2003cz}.

@@

\begin{acknowledgments}
  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work. \mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile) and F.~R. by the Fondecyt project No.~1160423. The Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
\end{acknowledgments}
%_-----------------------------------------------------------------------------------------

>>>>

The simplest possible extra term in the Lanczos--Lovelock action  is a quadratic construction of curvatures, called Gauss--Bonnet term, which reads
\begin{equation}\label{GB}
  \mathcal{L}_{\mathrm{GB}} = \dn{N}{x}\sqrt{-g}\left(\tilde{R}^2-4\tilde{R}_{\mu\nu}\tilde{R}^{\mu\nu}+\tilde{R}_{\alpha\beta\mu\nu}
  \tilde{R}^{\alpha\beta\mu\nu}\right),
\end{equation}
where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant. $\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively. In four dimensions, the Gauss--Bonnet term adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term. Nevertheless, it has been shown that this term can be relevant for conserved charges considerations in spacetimes with local AdS asymptotic~\cite{Aros:1999id}. Moreover, as it was reported in Ref.~\cite{Miskovic:2009bm}, the inclusion of topological invariants of the Euler class is equivalent to the program of Holographic Renormalization in the context of AdS/CFT. In dimensions higher than four it contributes to the field equations and it was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.

@@

In this work, we will focus on the region in the parameter space where
\begin{equation}
  \label{delta}
  \Delta\equiv\alpha_1^2-4\alpha_0\alpha_2\neq 0.
\end{equation}
This condition place us outside the Chern--Simons point. Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2016cs,*Troncoso:1999pk}. Exact solutions with torsion in Chern--Simons gravity are given in~\cite{Aros:2007nn,Banados:2003cz}. Its holographic properties in the Riemannian case where studied in Ref.~\cite{Banados:2005rz} and in \cite{Banados:2006fe} when torsion is considered.

@@

\begin{acknowledgments}
  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Olea, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work. \mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile) and F.~R. by the Fondecyt project No.~1160423. The Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
\end{acknowledgments}
%_-----------------------------------------------------------------------------------------
